\LARGE \bf Mean Estimation from One-Bit Measurements

### 2019-07-26 onebit_full.tex
removed 'bounded' from LAminimax result
corrected citation from Beran

@@

\node[int1, right of = source2, node distance = 1.2cm] (enc2) {Enc};

@@

\draw[->,line width = 2pt] (source2) -- (enc2); 
\node[int1, right of = enc2, node distance = 2.7cm ] (est) {Est};

@@

\node[right of = est, node distance = 1.2cm] (dest) {\small ${\theta}_n$};
%         \node [int1] (dec) [right of=dest, node distance = 1.5cm,  align=center] {\small Dec };
%\node [int1] (enc) [right of = dec, node distance = 3cm]{Enc}; 
%\draw[->,line width=2pt] (dec) -- (dest);
\draw[->, line width=1pt] (est) -- (dest);
\node at (2,-4.5) {(i) Centralized};
\end{tikzpicture}
%
\begin{tikzpicture}[node distance=2cm,auto]
  \node at (0,0) (source) {\small $X_1$} ;
  \node[int1, right of = source, node distance = 1.2cm] (enc1) {Enc 1};  
\draw[->,line width = 2pt] (source) -- (enc1);

 \node[below of = source, node distance = 1.5cm] (source2) {\small $X_2$};
\node[int1, right of = source2, node distance = 1.2cm] (enc2) {Enc 2};  
\draw[->,line width = 2pt] (source2) -- (enc2);

\node[below of = source2, node distance = 2.1cm] (source3) {\small $X_n$};
\node[int1, right of = source3, node distance = 1.2cm] (enc3) {Enc n};  
\draw[->,line width = 2pt] (source3) -- (enc3);

@@

\node[int1, right of = enc2, node distance = 3cm ] (est) {Est};
\draw[->,line width = 2pt] (enc1) -| node[above, xshift = -1cm] (mes1) {$B_1$} (est);   
\draw[->,line width = 2pt] (enc2) -- node[above, xshift = 0cm] (mes2) {$B_2$} (est);

@@

\node[right of = est, node distance = 1.2cm] (dest) {${\theta}_n$};
%         \node [int1] (dec) [right of=dest, node distance = 1.5cm,  align=center] {\small Dec };
%\node [int1] (enc) [right of = dec, node distance = 3cm]{Enc}; 
%\draw[->,line width=2pt] (dec) -- (dest);
\draw[->, line width=1pt] (est) -- (dest);
\node at (2,-4.5) {(ii) Adaptive};
\end{tikzpicture}
%
\begin{tikzpicture}[node distance=2cm,auto]
  \node at (0,0) (source) {$X_1$} ;
  \node[int1, right of = source, node distance = 1.2cm] (enc1) {Enc 1};  
\draw[->,line width = 2pt] (source) -- (enc1);

 \node[below of = source, node distance = 1.5cm] (source2) {$X_2$};
\node[int1, right of = source2, node distance = 1.2cm] (enc2) {Enc 2};  
\draw[->,line width = 2pt] (source2) -- (enc2);

\node[below of = source2, node distance = 2.1cm] (source3) {$X_n$};
\node[int1, right of = source3, node distance = 1.2cm] (enc3) {Enc n};  
\draw[->,line width = 2pt] (source3) -- (enc3);

@@

\node[int1, right of = enc2, node distance = 2.1cm ] (est) {Est};
\draw[->,line width = 2pt] (enc1) -| node[above, xshift = -1cm] (mes1) {$B_1$} (est);

@@

\subsection{Asymptotically optimal estimator}
Let $\left\{ \gamma_n,\, n\in \mathbb N \right\}$ be a strictly positive sequence. Consider the following estimator ${\theta}_n$ for $\theta$:  
\begin{equation}
\label{eq:sgd_alg}
\theta_n = \theta_{n-1} +  \gamma_n B_n, \quad n = 1,2,\ldots,
\end{equation}
where 
\begin{equation}
B_n = B_n (X_n,\theta_{n-1}) =\sgn (X_n - \theta_{n-1}).
\end{equation}
Define the $n$th step estimation as
\begin{equation} \label{eq:sgd_est}
\bar{\theta}_n =  \frac{1}{n} \sum_{i=1}^n  \theta_i. 
\end{equation}

We have the following results:
\begin{thm} \label{thm:sgd}
Consider the sequence $\left\{\bar{\theta}_n,\, n\in \mathbb N \right\}$ defined by \eqref{eq:sgd_est}. 
\begin{enumerate}
\item[(i)] Assume that $\left\{ \gamma_n,\, n\in \mathbb N \right\}$ satisfies
\begin{equation} \label{eq:conditions1}
\begin{cases}
\frac{\gamma_n - \gamma_{n+1}}{\gamma_n} = o(\gamma_n), &  \\
\sum_{n=1}^\infty \frac{\gamma_n^{(1+\lambda)/2}} {\sqrt{n}} < \infty, & 
\mathrm{for~some~}0< \lambda \leq 1
\end{cases}
\end{equation}
(e.g., $\gamma_n = n^{-\beta}$ for $\beta \in (0,1)$). Then
\[
\sqrt{n}\left( \bar{\theta}_n - \theta\right) \overset{d}{\to} \Ncal\left(0,\frac{1}{\eta(0)}\right).
\]
%
%
\item [(ii)] Assume in addition that $f(x)$ is continuously differentiable with finite Fisher information for location
\[
I_\theta = \int_{\mathbb R} \left( \frac{f'(x)}{f(x)}  \right)^2 f(x) dx < \infty.
\]
%(or that $f(x-\theta)$ is differentiable in quadratic mean).  
Then the estimator $\bar{\theta}_n$ is regular in the following sense: For $\theta \in \Theta$, $h\in \mathbb R$ and $n$ large enough such that $\theta+h/\sqrt{n} \in \Theta$, let $P^n_{\theta,h,n}$ be the measure with density $f(x-\theta - h/\sqrt{n})$. Then 
\begin{align}
\label{eq:sgd_part2}
\sqrt{n}\left( \bar{\theta}_n - \theta\right) \overset{d}{\to} \Ncal\left( h,\frac{1}{\eta(0)}\right),
\end{align}
under $P^n_{\theta,h,n}$. Consequently, for any bounded, symmetric, and quasi-convex function $L$, 
\begin{align*}
& \liminf_{c \to \infty} \liminf_{n \to \infty} \sup_{\tau\,:\,|\theta-\tau| \leq \frac{c}{\sqrt{n} }}  \mathbb E \left[ L\left( \sqrt{n}(\bar{\theta}_{n} - \tau) \right) \right]
\\
& \qquad \qquad \qquad \qquad \geq \mathbb E \left[ L (Z/\sqrt{\eta(0)}) \right],
\end{align*}
where $Z \sim \Ncal(0,1)$. 
%
%
\item[(iii)] Assume that, in addition to \eqref{eq:conditions1}, $\left\{ \gamma_n,\, n\in \mathbb N \right\}$ satisfies
\begin{equation} \label{eq:conditions2}
\begin{cases}
\gamma_n = o(n^{-2/3}), &  \\
\sum_{n=1}^\infty \gamma_n = \infty.  & \\
\end{cases}
\end{equation}
(e.g., $\gamma_n = n^{-\beta}$ with $\beta \in (2/3,1)$). Then
\[
 \ex{\left( \bar{\theta}_n - \theta \right)^2} = \frac{1}{n\eta(0)} + o(n^{-1}). 
\]
\end{enumerate}
\end{thm}

@@

Theorem~\ref{thm:sgd} implies that the estimator ${\theta}_n$, defined by \eqref{eq:sgd_est} and \eqref{eq:sgd_alg}, attains the maximal ARE as established by Theorem~\ref{thm:adpative_lower_bound}. %Furthermore, it is locally minimax in the sense that it attains the minimal MSE for all local alternatives of $\theta$. 
%
The update step \eqref{eq:sgd_alg} can be seen as a gradient descent step for the function $x\to |x|$ at the point $x=X_n - \theta_{n-1}$. Consequently, the procedure above is known as averaged stochastic gradient descent for minimizing $x \to |x|$ given the data $X_1,\ldots,X_n$. The minimal value of this optimization is the sample median, and Theorem~\ref{thm:sgd} provides conditions for the sequence of gradient steps so that the algorithm converges to this minimum. 
 \par
%Note that $\theta_0$ is not explicitly defined in equation \eqref{eq:sgd_est}. A reasonable initialization for $\theta_0$ is $\theta_0 = \mathbb E [\theta]$, although Theorem~\ref{thm:sgd} implies that the asymptotic behavior of the estimator is indifferent to this initialization. Thus, the optimal efficiency is attained regardless of the prior distribution on $\theta$ or the radius of the parameter space $\Theta$.  \\
In the encoding and estimating procedure \eqref{eq:sgd_alg} and \eqref{eq:sgd_est}, each one-bit message $B_n$ depends on its private sample as well as the current gradient descent estimate $\theta_{n-1}$. In this sense, each encoder in this algorithm interacts with previous one by using the current estimate.
%
As we explain next, 
it is possible to obtain the optimal efficiency of $\eta(0)\sigma^2$ with only one round of interactions among the encoders.

\begin{figure}
\begin{center}
\begin{tikzpicture}[node distance=2cm,auto]
  \node at (0,0) (source) {$X_1$} ;
  \node[int1, right of = source, node distance = 1.2cm] (enc1) {Enc};  
\draw[->,line width = 2pt] (source) -- (enc1);

@@

\node[below of = source, node distance = 1.7cm] (source3) {$X_{n_1}$};
\node[int1, right of = source3, node distance = 1.2cm] (enc3) {Enc};

@@

\node[int1, right of = enc3, node distance = 2.1cm ] (est) {Est 1};

@@

\node[below right = 0.75 and 1.5 of source3] (sourceB) {$X_{n_1 +1}$} ;
\node[int1, right of = sourceB, node distance = 1.7cm] (enc1B) {Enc$({\theta}_{n_1})$};  
\draw[->,line width = 2pt] (sourceB) -- (enc1B);

\node[below of = sourceB, node distance = 1.7cm] (source3B) {$X_n$};
\node[int1, right of = source3B, node distance = 1.7cm] (enc3B) {Enc$({\theta}_{n_1})$};  
\draw[->,line width = 2pt] (source3B) -- (enc3B); 
\node[below of = sourceB, node distance = 0.4cm] {$\vdots$};

\node[int1, right of = enc3B, node distance = 2.1cm ] (estB) {Est 2};

\draw[->,line width = 2pt] (enc1B) -| node[above, xshift = -0.8cm] {$B_{n_1+1}$} (estB);
\draw[->,line width = 2pt] (enc3B) -- node[above] {$B_n$} (estB);

@@

\draw[->,line width = 1pt] (est.east) -| +(1,-0.5) -- +(1,-2) -| (enc3B.north);

@@

\draw[->,line width = 1pt] (estB) -- +(0.8,0) node[right] {${\theta}_n$};
\end{tikzpicture}
\end{center}
\caption{Distributed encoding with a single interaction: The estimation obtained from the first $n_1$ bits in a distributed manner is to obtain another $n-n_1$ bits in a distributed manner. 
\label{fig:one_round}
}
\end{figure}

@@

\section{Distributed Estimation \label{sec:distributed}}
We now consider the distributed encoding setting in Figure~\ref{fig:setup}-(iii). In this setting each one-bit message $B_i$ is only a function of its private sample $X_i$, and hence $B_i$ characterized by its \emph{detection region}, defined as 
\[
A_i = \left\{ x \in \mathbb R \,:\, B_i(x) = 1 \right\}.
\]
Consequently, $B_i$ is of the form
\[
B_i = \begin{cases} 1 & X_i \in A_i, \\
-1 & X_i \notin A_i,
\end{cases} \quad i\in \mathbb N,
\]
where the detection region $A_i$ is a Borel set that is independent of $X_1,\ldots,X_n$.\\

@@

Theorem~\ref{thm:LAN1} provides conditions under which $B_1,\ldots,B_n$ defines a LAN family with a precision parameter given by the limit in \eqref{eq:LAN_lim}. 
%
Among these conditions, (iii) is arguably the strongest and hardest to verify. As we show in Section~\ref{subsec:threshold} below, this condition is satisfied, for example, when $A_1, \ldots,A_n$ are half lines whose starting points are drawn independently from some probability measure on $\mathbb R$. Similar ideas imply that condition (iii) holds whenever we choose the intervals consisting each $A_i$ according to some pre-specified distribution. \par
%
An important conclusion of  Theorem~\ref{thm:LAN1} follows from the local asymptotic minimax property of estimators in LAN models (e.g. \cite{van2000asymptotic}):
\begin{cor} \label{cor:LA_minimax}
Let ${\theta}_n$ be an estimator of $\theta \in \Theta$ from $B_1,\ldots,B_n$ with detection regions $A_1,\ldots,A_n$ such that conditions (i)-(iii) of Theorem~\ref{thm:LAN1} hold. Then for any bounded, symmetric, and quasi-convex function $L$, 
\begin{align*}
& \liminf_{c \to \infty} \liminf_{n \to \infty} \sup_{\tau\,:\,|\theta-\tau| \leq \frac{c}{\sqrt{n} }}  \mathbb E \left[ L\left( \sqrt{n}({\theta}_{n} - \tau) \right) \right]
\\
& \qquad \qquad \qquad \qquad \geq \mathbb E \left[ L (Z/\sqrt{\kappa(\theta)}) \right],
\end{align*}
where $Z \sim \Ncal(0,1)$. In particular, for $L(x) = x^2$,
\begin{align*}
\liminf_{c \to \infty} \liminf_{n \to \infty} \sup_{\tau\,:\,|\theta-\tau| \leq \frac{c}{\sqrt{n} }}  n \mathbb E  \left( {\theta}_{n} - \tau \right)^2 \geq 1/\kappa(\theta).
\end{align*}
\end{cor}
%
Corollary~\ref{cor:LA_minimax} says that when the messages define a LAN model, no estimator can attain MSE smaller than $1/\kappa(\theta)n + O(1/n)$ where $\kappa(\theta)$ is the precision parameter of the model at $\theta$. This fact poses the upper bound of $\kappa(\theta)\sigma^2$ for the ARE of estimators in such models. \par

@@

Figure~\ref{fig:minimax_support} illustrates an approximating to $\lambda^\star(dt)$ obtained by solving a discretized version of \eqref{eq:var_cvx_minimax} for the case when $f(x)$ is the normal density with variance $\sigma^2$. The minimal asymptotic risk $\kappa^\star(\theta)$ obtained this way is illustrated in Fig.~\ref{fig:minimax_ARE} as a function of the support size $T$. Also illustrated in these Figures is $\kappa_{\unif}$ which is the precision parameter corresponding to threshold values uniformly distribution over $\Theta = [-T,T]$, namely
\begin{align}
& \kappa_{\unif} \triangleq \min_{\theta \in [-T,T]} \frac{1}{2T}\int_{-T}^T \eta\left(t-\theta\right) dt \nonumber
 \\
& = 
\frac{1}{2T}\int_{-T}^{T} \eta\left(t\pm T\right) dt
= \frac{1}{2T}\int_{0}^{2T} \eta(t) dt  \label{eq:uniform_risk}. 
\end{align}
From Corollary~\ref{cor:LAN_thresh}, we conclude that the ARE under a uniform distribution is $\kappa_\unif \sigma^2$. \\

@@

\section{Conclusions \label{sec:conclusions}}
We considered the MSE risk in estimating the mean of a symmetric and log-concave distribution from a sequence of bits, where each bit is obtained by encoding a single sample from this distribution. 
%
In an adaptive encoding setting, we showed that no estimator can attain asymptotic relative efficiency (ARE) larger than that of the median of the samples. We also showed that this bound is tight by presenting two adaptive encoding and estimation procedures that are as efficient as the median. 
%We also characterized the one-step optimal scheme in this adaptive setting, i.e., the scheme that minimizes the risk given any set of previously obtained bits.
\par
In the distributed setting we provided conditions for local asymptotic normality of the encoded samples, which implies asymptotic minimax bound on both the risk and ARE. 
%
We conclude that under such conditions, the optimal estimation performance derived for the adaptive case can only be attained over a finite number of points, i.e., no scheme is uniformly optimal in the distributed setting. 
%
We further considered the special case of messages obtained by comparing against a prescribed sequence of thresholds. We characterized the performance of the optimal estimator from such messages using the density of these thresholds, and consider the threshold density that minimizes the minimax risk.
 \\

@@

\subsubsection*{Proof of (ii)}
The proof requires the following refinement of Theorem~\ref{thm:polyak_juditsky}.
\begin{thm} \label{thm:normal_expansion}
Set $\Delta_i = \theta_i - \theta$ and
$\bar{\Delta_i} = \frac{1}{n} \sum_{i=1}^n \Delta_i$. Assume that, in addition to Assumptions (i)-(iv) of Theorem~\ref{thm:polyak_juditsky}, there exists $K_1$ and $\lambda>0$ such that
\begin{equation}
\ex{\left| \varphi(Z_1) - \varphi(x+Z_1)  \right|} \leq K_1 |x|^{1+\lambda}
\label{eq:PJ_additional_cond}. 
\end{equation}
Then:
\begin{itemize}
\item[(i)] \begin{equation}
\sqrt{n} \bar{\Delta}_n = -\frac{1}{\sqrt{n}} \frac{1}{\psi'(0)} \sum_{i=1}^{n-1} \varphi(Z_i)+ o_{p.n}(1). \label{eq:normal_expansion_lem}
\end{equation}
where $o_{p,n}(1)$ converges in probability to $0$ as $n\to \infty$.
\item[(ii)] If $Z_1$ has continuously differentiable density $f(x)$ with finite location Fisher information 
\[
I_\theta = \int_{\mathbb R} \left( \frac{f'(x)}{f(x)} \right)^2 f(x) dx,
\]
then
\[
\sqrt{n} \left( \bar{\Delta}_n \right) \overset{d}{\to} \Ncal\left( \frac{-1}{\psi'(0)} \int_{\mathbb R} \varphi(x) f'(x) dx  , \frac{\chi(0)}{\psi'^2(0)} \right)
\]
under the local alternative $Z_1,\ldots,Z_n \sim \Prob^n_{h/\sqrt{n}}$ with density $f^n(x-h/\sqrt{n})$.
\end{itemize}
\end{thm}
The proof of Theorem~\ref{thm:normal_expansion} is given in Subsection~\ref{proof:thm:normal_expansion} below. \\

>>>>

\newcommand{\enc}{\mathsf{enc}}
\newcommand{\est}{\mathsf{est}}

@@

\node[int1, right of = source2, node distance = 1.2cm] (enc2) {$\enc$};

@@

\draw[->,line width = 2pt] (source2) -- (enc2); 
\node[int1, right of = enc2, node distance = 2.7cm ] (est) {$\est$};

@@

\node[right of = est, node distance = 1.2cm] (dest) {\small ${\theta}_n$};
%         \node [int1] (dec) [right of=dest, node distance = 1.5cm,  align=center] {\small Dec };
%\node [int1] (enc) [right of = dec, node distance = 3cm]{Enc}; 
%\draw[->,line width=2pt] (dec) -- (dest);
\draw[->, line width=1pt] (est) -- (dest);
\node at (2,-4.5) {(i) Centralized};
\end{tikzpicture}
%
\begin{tikzpicture}[node distance=2cm,auto]
  \node at (0,0) (source) {\small $X_1$} ;
  \node[int1, right of = source, node distance = 1.2cm] (enc1) {$\enc$};  
\draw[->,line width = 2pt] (source) -- (enc1);

 \node[below of = source, node distance = 1.5cm] (source2) {\small $X_2$};
\node[int1, right of = source2, node distance = 1.2cm] (enc2) {$\enc$};  
\draw[->,line width = 2pt] (source2) -- (enc2);

\node[below of = source2, node distance = 2.1cm] (source3) {\small $X_n$};
\node[int1, right of = source3, node distance = 1.2cm] (enc3) {$\enc$};  
\draw[->,line width = 2pt] (source3) -- (enc3);

@@

\node[int1, right of = enc2, node distance = 3cm ] (est) {$\est$};
\draw[->,line width = 2pt] (enc1) -| node[above, xshift = -1cm] (mes1) {$B_1$} (est);   
\draw[->,line width = 2pt] (enc2) -- node[above, xshift = 0cm] (mes2) {$B_2$} (est);

@@

\node[right of = est, node distance = 1.2cm] (dest) {${\theta}_n$};
%         \node [int1] (dec) [right of=dest, node distance = 1.5cm,  align=center] {\small Dec };
%\node [int1] (enc) [right of = dec, node distance = 3cm]{Enc}; 
%\draw[->,line width=2pt] (dec) -- (dest);
\draw[->, line width=1pt] (est) -- (dest);
\node at (2,-4.5) {(ii) Adaptive};
\end{tikzpicture}
%
\begin{tikzpicture}[node distance=2cm,auto]
  \node at (0,0) (source) {$X_1$} ;
  \node[int1, right of = source, node distance = 1.2cm] (enc1) {$\enc$};  
\draw[->,line width = 2pt] (source) -- (enc1);

 \node[below of = source, node distance = 1.5cm] (source2) {$X_2$};
\node[int1, right of = source2, node distance = 1.2cm] (enc2) {$\enc$};  
\draw[->,line width = 2pt] (source2) -- (enc2);

\node[below of = source2, node distance = 2.1cm] (source3) {$X_n$};
\node[int1, right of = source3, node distance = 1.2cm] (enc3) {$\enc$};  
\draw[->,line width = 2pt] (source3) -- (enc3);

@@

\node[int1, right of = enc2, node distance = 2.1cm ] (est) {$\est$};
\draw[->,line width = 2pt] (enc1) -| node[above, xshift = -1cm] (mes1) {$B_1$} (est);

@@

\subsection{Asymptotically optimal estimator}
Let $\left\{ \gamma_n \right\}_{n\in \mathbb N}$ be a strictly positive sequence. Consider the following estimator ${\theta}_n$ for $\theta$:  
\begin{equation}
\label{eq:sgd_alg}
\theta_n = \theta_{n-1} +  \gamma_n B_n, \quad n = 1,2,\ldots,
\end{equation}
where 
\begin{equation}
B_n = B_n (X_n,\theta_{n-1}) =\sgn (X_n - \theta_{n-1}).
\end{equation}
Define the $n$th step estimation as
\begin{equation} \label{eq:sgd_est}
\bar{\theta}_n =  \frac{1}{n} \sum_{i=1}^n  \theta_i. 
\end{equation}

We have the following results:
\begin{thm} \label{thm:sgd}
Consider the sequence $\left\{\bar{\theta}_n \right\}_{n\in \mathbb N}$ defined by \eqref{eq:sgd_est}. 
\begin{enumerate}
\item[(i)] Assume that $\left\{ \gamma_n \right\}_{n\in \mathbb N}$ satisfies
\begin{equation} \label{eq:conditions1}
\begin{cases}
\frac{\gamma_n - \gamma_{n+1}}{\gamma_n} = o(\gamma_n), &  \\
\sum_{n=1}^\infty \frac{\gamma_n^{(1+\lambda)/2}} {\sqrt{n}} < \infty, & 
\mathrm{for~some~}0< \lambda \leq 1
\end{cases}
\end{equation}
(e.g., $\gamma_n = n^{-\beta}$ for $\beta \in (0,1)$). Then
\[
\sqrt{n}\left( \bar{\theta}_n - \theta\right) \overset{d}{\to} \Ncal\left(0,\frac{1}{\eta(0)}\right).
\]
%
%
\item [(ii)] Assume in addition that $f(x)$ is continuously differentiable with finite Fisher information for location
\[
I_\theta = \int_{\mathbb R} \left( \frac{f'(x)}{f(x)}  \right)^2 f(x) dx < \infty.
\]
%(or that $f(x-\theta)$ is differentiable in quadratic mean).  
Then for any bounded, symmetric, and quasi-convex function $L$, 
\begin{align} 
& \liminf_{c \to \infty} \liminf_{n \to \infty} \sup_{\tau\,:\,|\theta-\tau| \leq \frac{c}{\sqrt{n} }}  \mathbb E \left[ L\left( \sqrt{n}(\bar{\theta}_{n} - \tau) \right) \right] \nonumber 
\\
& \qquad \qquad \qquad \qquad = \mathbb E \left[ L (Z/\sqrt{\eta(0)}) \right], \label{eq:attaining_LAM}
\end{align}
where $Z \sim \Ncal(0,1)$. 
%
%
\item[(iii)] Assume that, in addition to \eqref{eq:conditions1}, $\left\{ \gamma_n \right\}_{n\in \mathbb N}$ satisfies
\begin{equation} \label{eq:conditions2}
\begin{cases}
\gamma_n = o(n^{-2/3}), &  \\
\sum_{n=1}^\infty \gamma_n = \infty.  & \\
\end{cases}
\end{equation}
(e.g., $\gamma_n = n^{-\beta}$ with $\beta \in (2/3,1)$). Then
\[
 \ex{\left( \bar{\theta}_n - \theta \right)^2} = \frac{1}{n\eta(0)} + o(n^{-1}). 
\]
\end{enumerate}
\end{thm}

@@

Theorem~\ref{thm:sgd} implies that the estimator ${\theta}_n$, defined by \eqref{eq:sgd_est} and \eqref{eq:sgd_alg}, attains the maximal ARE as established by Theorem~\ref{thm:adpative_lower_bound}. %Furthermore, it is locally minimax in the sense that it attains the minimal MSE for all local alternatives of $\theta$. 
%
The update step \eqref{eq:sgd_alg} can be seen as a gradient descent step for the function $x\to |x|$ at the point $x=X_n - \theta_{n-1}$. Consequently, the procedure above is known as averaged stochastic gradient descent for minimizing $x \to |x|$ given the data $X_1,\ldots,X_n$. The minimal value of this optimization is the sample median, and Theorem~\ref{thm:sgd} provides conditions for the sequence of gradient steps so that the algorithm converges to this minimum. In particular, parts (i) and (ii) of Theorem~\ref{thm:sgd} provides conditions under which $\bar{\theta}$ is asymptotically normal and local asymptotically minimax, in the sense that it attains the local asymptotic minimax bound \cite{van2000asymptotic}. Part (iii) provides an additional condition under which $\bar{\theta}$ also converges in its second moment. 
 \par
%Note that $\theta_0$ is not explicitly defined in equation \eqref{eq:sgd_est}. A reasonable initialization for $\theta_0$ is $\theta_0 = \mathbb E [\theta]$, although Theorem~\ref{thm:sgd} implies that the asymptotic behavior of the estimator is indifferent to this initialization. Thus, the optimal efficiency is attained regardless of the prior distribution on $\theta$ or the radius of the parameter space $\Theta$.  \\
In the encoding and estimating procedure \eqref{eq:sgd_alg} and \eqref{eq:sgd_est}, each one-bit message $B_n$ depends on its private sample as well as the current gradient descent estimate $\theta_{n-1}$. In this sense, each encoder in this algorithm interacts with previous one by using the current estimate.
%
As we explain next, 
it is possible to obtain the optimal efficiency of $\eta(0)\sigma^2$ with only one round of interactions among the encoders.

\begin{figure}
\begin{center}
\begin{tikzpicture}[node distance=2cm,auto]
  \node at (0,0) (source) {$X_1$} ;
  \node[int1, right of = source, node distance = 1.2cm] (enc1) {$\enc$};  
\draw[->,line width = 2pt] (source) -- (enc1);

@@

\node[below of = source, node distance = 1.7cm] (source3) {$X_{n_1}$};
\node[int1, right of = source3, node distance = 1.2cm] (enc3) {$\enc$};

@@

\node[int1, right of = enc3, node distance = 2.1cm ] (est) {$\est_1$};

@@

\node[below right = 0.75 and 1.5 of source3] (sourceB) {$X_{n_1 +1}$} ;
\node[int1, right of = sourceB, node distance = 1.7cm] (enc1B) {$\enc$};  
\draw[->,line width = 2pt] (sourceB) -- (enc1B);

\node[below of = sourceB, node distance = 1.7cm] (source3B) {$X_n$};
\node[int1, right of = source3B, node distance = 1.7cm] (enc3B) {$\enc$};  
\draw[->,line width = 2pt] (source3B) -- (enc3B); 
\node[below of = sourceB, node distance = 0.4cm] {$\vdots$};

\node[int1, right of = enc3B, node distance = 2.1cm ] (estB) {$\est_2$};

\draw[->,line width = 2pt] (enc1B) -| node[above, xshift = -0.5cm] {$B_{n_1+1}$} (estB);
\draw[->,line width = 2pt] (enc3B) -- node[above] {$B_n$} (estB);

@@

\draw[->,line width = 0.5pt] (est.east) -| +(1.3,-0.5) -- +(1.3,-2.5) -| (enc3B.north);

@@

\draw[->,line width = 0.5pt] (estB) -- +(0.8,0) node[right] {${\theta}_n$};
\node[below of = enc1B, node distance = 0.5cm] {$\vdots$};

\end{tikzpicture}
\end{center}
\caption{Distributed encoding with a single interaction: The estimation obtained from the first $n_1$ bits in a distributed manner is to obtain another $n-n_1$ bits in a distributed manner. 
\label{fig:one_round}
}
\end{figure}

@@

\section{Distributed Estimation \label{sec:distributed}}
We now consider the distributed encoding setting in Figure~\ref{fig:setup}-(iii). In this setting each one-bit message $B_i$ is only a function of its private sample $X_i$, and hence $B_i$ characterized by its \emph{detection region}, defined as 
\[
A_i = \left\{ x \in \mathbb R \,:\, B_i(x) = 1 \right\}.
\]
Consequently, $B_i$ is of the form
\[
B_i = \begin{cases} 1 & X_i \in A_i, \\
-1 & X_i \notin A_i,
\end{cases} \quad i\in \mathbb N,
\]
where the detection region $A_i$ is a Borel set that is independent of $X_1,\ldots,X_n$.\par

@@

Theorem~\ref{thm:LAN1} provides conditions under which $B_1,\ldots,B_n$ defines a LAN family with a precision parameter given by the limit in \eqref{eq:LAN_lim}. 
%
Among these conditions, (iii) is arguably the strongest and hardest to verify. As we show in Section~\ref{subsec:threshold} below, this condition is satisfied, for example, when $A_1, \ldots,A_n$ are half lines whose starting points are drawn independently from some probability measure on $\mathbb R$. Similar ideas imply that condition (iii) holds whenever we choose the intervals consisting each $A_i$ according to some pre-specified distribution. \par
%
An important conclusion of  Theorem~\ref{thm:LAN1} follows from the local asymptotic minimax property of estimators in LAN models \cite{van2000asymptotic}):
\begin{cor} \label{cor:LA_minimax}
Let ${\theta}_n$ be an estimator of $\theta \in \Theta$ from $B_1,\ldots,B_n$ with detection regions $A_1,\ldots,A_n$ such that conditions (i)-(iii) of Theorem~\ref{thm:LAN1} hold. Then for any symmetric and quasi-convex function $L$, 
\begin{align*}
& \liminf_{c \to \infty} \liminf_{n \to \infty} \sup_{\tau\,:\,|\theta-\tau| \leq \frac{c}{\sqrt{n} }}  \mathbb E \left[ L\left( \sqrt{n}({\theta}_{n} - \tau) \right) \right]
\\
& \qquad \qquad \qquad \qquad \geq \mathbb E \left[ L (Z/\sqrt{\kappa(\theta)}) \right],
\end{align*}
where $Z \sim \Ncal(0,1)$. In particular, for $L(x) = x^2$,
\begin{align*}
\liminf_{c \to \infty} \liminf_{n \to \infty} \sup_{\tau\,:\,|\theta-\tau| \leq \frac{c}{\sqrt{n} }}  n \mathbb E  \left( {\theta}_{n} - \tau \right)^2 \geq 1/\kappa(\theta).
\end{align*}
\end{cor}
%
Corollary~\ref{cor:LA_minimax} says that when the messages define a LAN model, no estimator can attain MSE smaller than $1/\kappa(\theta)n + O(1/n)$ where $\kappa(\theta)$ is the precision parameter of the model at $\theta$. This fact poses the upper bound of $\kappa(\theta)\sigma^2$ for the ARE of estimators in such models. \par

@@

Figure~\ref{fig:minimax_support} illustrates an approximating to $\lambda^\star(dt)$ obtained by solving a discretized version of \eqref{eq:var_cvx_minimax} for the case when $f(x)$ is the normal density with variance $\sigma^2$. The minimal asymptotic risk $\kappa^\star(\theta)$ obtained this way is illustrated in Fig.~\ref{fig:minimax_ARE} as a function of the support size $T$. Also illustrated in these Figures is $\kappa_{\unif}$ which is the precision parameter corresponding to threshold values uniformly distribution over $\Theta = [-T,T]$, namely
\begin{align}
& \kappa_{\unif} \triangleq \min_{\theta \in [-T,T]} \frac{1}{2T}\int_{-T}^T \eta\left(t-\theta\right) dt \nonumber
 \\
& = 
\frac{1}{2T}\int_{-T}^{T} \eta\left(t\pm T\right) dt
= \frac{1}{2T}\int_{0}^{2T} \eta(t) dt  \label{eq:uniform_risk}. 
\end{align}
From Corollary~\ref{cor:LAN_thresh}, we conclude that the ARE under a uniform distribution is $\kappa_\unif \sigma^2$. \par

@@

\section{Conclusions \label{sec:conclusions}}
We considered the MSE risk in estimating the mean of a symmetric and log-concave distribution from a sequence of bits, where each bit is obtained by encoding a single sample from this distribution. 
%
In an adaptive encoding setting, we showed that no estimator can attain asymptotic relative efficiency (ARE) larger than that of the median of the samples. We also showed that this bound is tight by presenting two adaptive encoding and estimation procedures that are as efficient as the median. 
%We also characterized the one-step optimal scheme in this adaptive setting, i.e., the scheme that minimizes the risk given any set of previously obtained bits.
\par
In the distributed setting we provided conditions for local asymptotic normality of the encoded samples, which implies asymptotic minimax bound on both the risk and ARE. 
%
We conclude that under such conditions, the optimal estimation performance derived for the adaptive case can only be attained over a finite number of points, i.e., no scheme is uniformly optimal in the distributed setting. 
%
We further considered the special case of messages obtained by comparing against a prescribed sequence of thresholds. We characterized the performance of the optimal estimator from such messages using the density of these thresholds, and consider the threshold density that minimizes the minimax risk.

@@

\subsubsection*{Proof of (ii)}
We first show that the estimator $\bar{\theta}_n$ is regular in the following sense: For $\theta \in \Theta$, $h\in \mathbb R$ and $n$ large enough such that $\theta+h/\sqrt{n} \in \Theta$, let $\Prob_{\theta,n}$ be a product probability measure on $\mathbb R^n$ with density $f(x-\theta - h/\sqrt{n})$ in each of its $n$ coordinates. Then 
\begin{align}
\label{eq:sgd_part2}
\sqrt{n}\left( \bar{\theta}_n - \theta\right) \overset{d}{\to} \Ncal\left( h,\frac{1}{\eta(0)}\right),
\end{align}
under $\Prob_{\theta,n}$. Under the assumption that $f(x)$ is continuously differentiable with finite Fisher information for location, the model $\{Z_n+ \theta\}_{n \in \mathbb N}$ is differentiable in quadratic mean \cite[Exm. 7.8]{van2000asymptotic} and hence local asymptotically normal (LAN) in the sense that
\[
\log \left(\frac{\Prob_{\theta,n}(X^n)}{\Prob_{\theta} (X^n) }\right) = h \eta^{-1/2}(0) Z - \frac{1}{2} h^2 \eta^{-1}(0)  + o_{p,n}(1),
\]
where $Z\sim \Ncal(0,1)$ and $o_{p,n}(1) \to 0$ as $n\to \infty$ under $\Prob_{\theta}$. 
 The random variable $L\left(\sqrt{n} \left(\bar{\theta}_n - \tau \right) \right)
$ is bounded, and hence 
\[
\ex{ L\left(\sqrt{n} \left(\bar{\theta}_n - \tau \right) \right)} \to \ex{L\left(\sqrt{n} \left(\bar{\theta}_n - \tau \right) \right)}
\]

Any regular estimators in a LAN model satisfies \eqref{eq:attaining_LAM} \cite{beran1995role}, hence the proof of (ii) follows once we establish the regularity of $\bar{\theta}_n$. \\

%Any estimator satisfying \eqref{eq:sgd_part2} also satisfies \eqref{eq:attaining_LAM} \cite[P. 157]{ibragimov2013statistical}.
